Towards a standardized evaluation of imputation methodology

### 2022-09-16 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex

[TODO: make this section less stringent and focus more on the suitability of the performancxe evaluation in the context of the study aim.]

@@

(i) The estimates should preferably be unbiased. Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results. In most cases, unbiased estimates may be expected under a MCAR missingness mechanism.

@@

(iv) Predictive accuracy may be quantified using the root mean squared error (RMSE). We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values \citep[][\S 2.5]{buur18}. However, if a study is aimed at obtaining predictions or classification and inferential validity is not of interest, the RMSE of the predictions may yield valuable information about the methods' efficiency in terms of both accuracy and precision. [TODO: make this an optional bullet point for prediction only: out of sample mean squared prediction error can be compared between methods.]

[TODO: intrgrate this paragraph into the above!] Performance measures should express how well the obtained estimates approximate the comparative truth, but not every metric is suited for the evaluation of imputation methodology. For example, using measures of predictive accuracy or the RMSE (root mean squared error) as the evaluation criterion may increase the rate of false positives \citep[][\S 2.6]{buur18}. And, under certain simulation conditions, an otherwise useful metric such as relative bias can distort the interpretation of results (i.e., a negligible absolute bias may yield an infinite relative bias if the true value of the estimand is zero). The usefulness of evaluation criteria may thus depend on the simulation goal and estimand.

If the goal is inference--not prediction--then the uncertainty about estimates needs to be properly quantified. To capture this uncertainty, the standard errors of the estimates should be correctly calculated, which requires multiple imputation. The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied. After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p. 591]{neym34}. [TODO: add nuance about coverage rate > 95\% irt CIW and efficiency.]

The choice of performance measures may inadvertently distort statistical properties of the imputed data. Developers often only inquire about the `accuracy' (i.e. how well can the method reproduce the original data). Merely focusing on the discrepancy between the true and imputed data would open the door for invalid inferences.

>>>>

[TODO: make this section less stringent and focus more on the suitability of the performance evaluation in the context of the study aim. The take-away should be that simulation results cannot be easily compared if the performance measures differ. So we suggest to use CCA as benchmark method, and if possible also complete-data inference, to provide a lower and upper limit for expected imputation method performance. And even if the goal of the simulation effort is imputation method comparison, one should not just evaluate relative performance between imputation methods, but also report statistical properties of estimates such as bias, confidence intervals, and the coverage rate of the confidence intervals.]

@@

(i) The estimates should preferably be unbiased. Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results (i.e., a negligible absolute bias may yield an infinite relative bias if the true value of the estimand is zero). In most cases, unbiased estimates may be expected under a MCAR missingness mechanism.

@@

(iv) Predictive accuracy may be quantified using the root mean squared error (RMSE). We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may increase the rate of false positives \citep[][]{buur18}. However, if a study is aimed at obtaining predictions, and inferential validity is not of interest, the RMSE of the predicted values can be compared between method, which may yield valuable information about the methods' efficiency in terms of both accuracy and precision.

If the goal is inference--not prediction--then the uncertainty about estimates needs to be properly quantified. To capture this uncertainty, the standard errors of the estimates should be correctly calculated, which requires multiple imputation. The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied.
